### 2016-03-03 tex/refactor.tex
Updates from ShareLaTeX

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|l|c|c|c|c}
Mostly from Fowler'99& & & & & Developer\\
 (with a few additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey\\\hline
  Alt. Classes with Diff. Interfaces & & & & + & \\
  Combinatorial Explosion~\cite{Kerievsky2004} & & & & + & \\
  Comments & & & 11 & + & VL\\
  Conditional Complexity~\cite{Kerievsky2004} & & & 14 & + & ?\\
  Data Class & Data Class & & & + &\\
  Data Clumps &  &  & & + &\\
  Divergent Change & & & & + & \\
  Duplicated Code & Significant Duplication\textsuperscript{*} & + & 1 & + & VH\\
  Feature Envy & Feature Envy & & 8 & + &\\

  Inappropriate Intimacy & & - & & + & L\\
  Indecent Exposure~\cite{Kerievsky2004} & & & & + & ?\\
  Incomplete Library Class & & & & + &\\
  Large Class & God Class/Brain Class & ++ & 4 & + & VH\\
  Lazy Class/Freeloader & & - & 7 & + &\\
  Long Method & Brain Method & ++ & 2 & + & VH\\
  Long Parameter List &  & + & 9 & + & L \\

  Message Chains & & & & + & H\\
  Middle Man & &  & & + &\\
  Oddball Solution~\cite{Kerievsky2004} & & & & + & \\
  Parallel Inheritance Hierarchies & & & & + &\\
  Primitive Obsession &  & & & + &\\

  Refused Bequest & Refused Parent Bequest & - & & + & \\ 
  Shotgun Surgery & Shotgun Surgery & & & + & \\
  Solution Sprawl~\cite{Kerievsky2004} & & & & + &\\
  Speculative Generality & & & & + & L\\
  Switch Statements &  & - & & & L\\
  Temporary Field & & - & & + & ?\\
  \end{tabular}
%   \multicolumn{5}{l}{} \\
%   \multicolumn{5}{l}{\textsuperscript{\textdagger} + Exact rule, - Possibly related rule} \\
%     \multicolumn{5}{l}{\textsuperscript{\textdaggerdbl} + is included in list} \\
%     \multicolumn{5}{l}{\textsuperscript{\textexclamdown} VH: Very High, H: High, L: Low, VL: Very Low, ?: Still investigating} \\
%   \multicolumn{5}{l}{\textsuperscript{*} Computation is based on more than simple metrics.} \\

\caption{List of bad code smells from different sources. Empty cells
denote some bad smell proposed by Fowler'99 that was not found relevant
in subsequent studies. Note that there are many blank cells.}
\label{fig:smells}
\end{figure*}