### 2016-03-08 tex/refactor.tex
Updates from ShareLaTeX

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|c|c|c|c|c}
Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
 (with a some additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
  Comments & & & 11 & \checkmark & VL\\
  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
  Data Class & \checkmark & & & \checkmark &\\
  Data Clumps &  &  & & \checkmark &\\
  Divergent Change & & & & \checkmark & \\
  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
  Feature Envy & \checkmark & & 8 & \checkmark &\\

  Inappropriate Intimacy & & \checkmark & & \checkmark & L\\
  Indecent Exposure~\cite{Kerievsky2005} & & & & \checkmark & ?\\
  Incomplete Library Class & & & & \checkmark &\\
  Large Class & \checkmark & \checkmark & 4 & \checkmark & VH\\
  Lazy Class/Freeloader & & \checkmark & 7 & \checkmark &\\
  Long Method & \checkmark& \checkmark & 2 & \checkmark & VH\\
  Long Parameter List &  & \checkmark & 9 & \checkmark & L \\

  Message Chains & & & & \checkmark & H\\
  Middle Man & &  & & \checkmark &\\
  Oddball Solution~\cite{Kerievsky2005} & & & & \checkmark & \\
  Parallel Inheritance Hierarchies & & & & \checkmark &\\
  Primitive Obsession &  & & & \checkmark &\\

  Refused Bequest & \checkmark & \checkmark & & \checkmark & \\ 
  Shotgun Surgery & \checkmark& & & \checkmark & \\
  Solution Sprawl~\cite{Kerievsky2005} & & & & \checkmark &\\
  Speculative Generality & & & & \checkmark & L\\
  Switch Statements &  & \checkmark & & & L\\
  Temporary Field & & \checkmark & & \checkmark & ?\\
  \end{tabular}
%   \multicolumn{5}{l}{} \\
%   \multicolumn{5}{l}{\textsuperscript{\textdagger} + Exact rule, - Possibly related rule} \\
%     \multicolumn{5}{l}{\textsuperscript{\textdaggerdbl} + is included in list} \\
%     \multicolumn{5}{l}{\textsuperscript{\textexclamdown} VH: Very High, H: High, L: Low, VL: Very Low, ?: Still investigating} \\
%   \multicolumn{5}{l}{\textsuperscript{*} Computation is based on more than simple metrics.} \\

\caption{List of bad code smells from different sources.  Check marks (\checkmark) denote that a bad smell was mentioned.
Numbers of symbolic labels (e.g. "VH") denote  a priorization comment.
``?'' indicates lack of consensus on some bad smell. Empty cells
denote some bad smell listed in column one that was not found relevant
in subsequent studies.
Note that there are many blank cells.}
\label{fig:smells}
\end{figure*}