### 2016-03-11 tex/rq2.tex

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\newcommand{\ZZ}{.}
\begin{figure*}
\renewcommand{\baselinestretch}{0.8} 
\scriptsize  
\centering

\label{my-label}

\begin{tabular}{c|rrrr|rrrr|rrrr|rrrr|rrrr}
\multicolumn{1}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} }} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Ant}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Ivy}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Lucene}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Jedit}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Poi}} \\
\multicolumn{1}{c@{}}{\multirow{-2}{*}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Features}}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} \\
wmc &\ZZ& 92 & 100 & 100 &\ZZ& 95 & 100 & 100 & 14 & 95 & 100 &\ZZ&\ZZ& 63 &\ZZ&\ZZ&\ZZ& 100 & 100 &\ZZ\\
dit &\ZZ& 77 & 100 &\ZZ&\ZZ& 87 & 100 &\ZZ&\ZZ& 80 & 100 &\ZZ&\ZZ& 72 & 100 & 100 &\ZZ& 46 & 100 &\ZZ\\
noc &\ZZ& 20 & 100 &\ZZ&\ZZ&\ZZ& 100 &\ZZ&\ZZ& 26 &\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ\\
cbo & 83 & 99 & 100 & 100 & 65 & 100 & 100 & 100 & 13 & 94 & 100 & 100 &\ZZ& 100 & 100 &\ZZ&\ZZ& 74 & 100 &\ZZ\\
rfc & 100 & 100 & 100 &\ZZ& 35 & 95 & 100 &\ZZ& 100 & 83 & 100 &\ZZ& 100 & 100 & 100 & 100 & 100 & 95 & 100 &\ZZ\\
lcom &\ZZ& 98 & 100 & 100 &\ZZ& 100 & 100 & 100 &\ZZ& 94 & 100 &\ZZ&\ZZ& 100 & 100 &\ZZ&\ZZ& 100 & 100 & 100 \\
ca &\ZZ& 93 & 100 &\ZZ&\ZZ& 95 & 100 &\ZZ& 87 & 89 & 100 &\ZZ&\ZZ& 63 & 100 & 100 &\ZZ& 74 & 100 &\ZZ\\
ce &\ZZ& 100 & 100 &\ZZ&\ZZ& 97 & 100 &\ZZ&\ZZ& 90 & 100 &\ZZ&\ZZ& 100 & 100 & 100 &\ZZ& 64 & 100 &\ZZ\\
npm &\ZZ& 88 & 100 &\ZZ& 35 & 97 & 100 &\ZZ&\ZZ& 93 & 100 & 100 &\ZZ& 100 & 100 &\ZZ&\ZZ& 100 & 100 &\ZZ\\
lcom3 &\ZZ& 90 & 100 &\ZZ&\ZZ& 95 & 100 &\ZZ& 13 & 79 & 100 & 100 &\ZZ& 63 & 100 & 100 &\ZZ& 92 & 100 & 100 \\
loc & 100 & 99 & 100 & 100 & 65 & 97 & 100 & 100 &\ZZ& 100 & 100 & 100 & 100 & 100 &\ZZ& 100 & 100 & 100 & 100 & 100 \\
dam &\ZZ& 21 & 100 &\ZZ&\ZZ& 22 & 100 &\ZZ&\ZZ& 55 & 100 &\ZZ&\ZZ& 45 & 100 & 100 &\ZZ& 73 & 100 &\ZZ\\
moa &\ZZ& 67 & 100 &\ZZ&\ZZ& 82 & 100 &\ZZ&\ZZ& 60 & 100 & 100 &\ZZ& 54 & 100 & 100 &\ZZ& 58 & 100 &\ZZ\\
mfa & 17 & 93 & 100 &\ZZ&\ZZ& 90 & 100 &\ZZ&\ZZ& 80 & 100 &\ZZ&\ZZ& 72 & 100 &\ZZ&\ZZ& 72 & 100 &\ZZ\\
cam &\ZZ& 99 & 100 & 100 & 100 & 100 & 100 & 100 &\ZZ& 94 & 100 &\ZZ&\ZZ& 100 & 100 &\ZZ&\ZZ& 98 & 100 & 100 \\
ic &\ZZ& 52 & 100 & 100 &\ZZ& 70 &\ZZ&\ZZ&\ZZ& 68 & 100 &\ZZ&\ZZ& 36 & 100 &\ZZ&\ZZ& 43 & 100 & 100 \\
cbm &\ZZ& 59 & 100 &\ZZ&\ZZ& 85 &\ZZ&\ZZ&\ZZ& 71 & 100 &\ZZ&\ZZ& 36 & 100 & 100 &\ZZ& 67 & 100 &\ZZ\\
amc &\ZZ& 99 &\ZZ&\ZZ&\ZZ& 95 & 100 &\ZZ& 13 & 100 &\ZZ&\ZZ&\ZZ& 100 & 100 & 100 &\ZZ& 97 &\ZZ&\ZZ\\
max cc &\ZZ& 87 & 100 & 100 &\ZZ& 85 & 100 & 100 &\ZZ& 71 &\ZZ&\ZZ&\ZZ& 45 & 100 &\ZZ&\ZZ& 63 & 100 &\ZZ\\
avg cc & 17 & 99 & 100 & 100 &\ZZ& 95 & 100 & 100 & 13 & 98 &\ZZ&\ZZ&\ZZ& 100 & 100 & 100 &\ZZ& 92 & 100 & \ZZ\\\hline
\end{tabular}
\caption{Percentage counts of  how often an approach recommends changing a code metric
(in 40 runs). Cells marked with ``.'' indicate  0\%. For the Shatnami and Alves et al.
columns,  metrics are not recommended if they fail the $p \le 0.05$ test of \tion{p}.
For CD, cells are blank when two centroids have the same value for the same code
metrics. For XTREE, cells are blanks when they do not appear in the delta
between branches.  The most important feature of this data is that in the XTREE column, the changes occur to only a small number of code metrics. That is, all the other methods are far more {\em verbose} than XTREE (where ``verbose'' is defined in \tion{disc}).}\label{fig:counts}
\end{figure*}